---
title: "Exploring the Credit Card data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
```

### Load data

<!-- Make sure your data and R Markdown files are in the same directory. When loaded -->
<!-- your data file will be called `brfss2013`. Delete this note when before you submit  -->
<!-- your work.  -->

```{r load-data}
df=read.csv("creditBalance.csv") 
```

* * *

## Part 1: Data
```{r data}
df = df %>%
  mutate(bal_to_limit_ratio=Balance/Limit)
df.new=select(df,-Gender,-Student,-Married)
# df.new1 = df.new %>%
#   mutate(bal_to_limit_ratio=Balance/Limit)
```

## Part 2: Questions to address

**Quesion 1:** What does the data and column say?

**Quesion 2:**Can we find relation between different columns?

**Quesion 3:** Interprete using graphical tool.


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Quesion 1:**


```{r}
dim(df)
names(df)
```
So the data set called "creditBalance" consists of 310 rows and 11 columns named "Income"    "Limit"     "Rating"    "Cards"     "Age"       "Education" "Gender"    "Student"   "Married" "bal_to_limit_ratio"  and "Balance" where Gender, Student, Married are nominal categorical and we can think Cards, education as ordinal variables(Although I will do some numerical analysis on cards) and the rest are numerical. 
```{r}
summary(df.new)
```
So what we can realize is that the people usually start having credit cards at the age of 24. Their income is 50(I think it is measured in 1000 but I might be wrong or missed the description somewhere) and limit is 5485, rating is 405 and balance is 671 on average. The balance to limit ratio varies mostly from 7 to 15 percent.

**RQuesion 2:**
```{r}
library(dplyr)
table(df$Gender,df$Student)
```
We can go a liitle ahead;
```{r}
library(dplyr)
df %>% group_by(., Gender,Student) %>% 
 dplyr:: select(., Gender, Student, Income, Balance) %>% 
  summarise(., Mean_Inc=mean(Income), Mean_Bal=mean(Balance),n())
```

We have 130 male non-student, 15 male student, 141 female non-student and 24 female student. We can see that if a user is student they have less income but more balance. Usually the male students have a significant amount of high balance 1004.53.
```{r}
df %>% group_by(., Married) %>% 
  select(., Married, Income, Balance) %>% 
  summarise(., Mean_Inc=mean(Income), Mean_Bal=mean(Balance),n())
```
We have 192 married user and 118 unmarried. Moreover, the married user earns more and keep balance lower than the other.

```{r}
library(dplyr)
df1=df %>% group_by(., Education) %>% 
  select(., Education, Income, Balance) %>% 
  summarise(., Mean_Inc=mean(Income), Mean_Bal=mean(Balance),number=n())
  df1[order(df1$number),]
```

So we can see that the most of the user is around education level 16.



**Research quesion 3:**

```{r}
plot(df$bal_to_limit_ratio)
```
So we can see that it is a completely scattered diagram




```{r}


library(dplyr)
library(ggplot2)
data(df)

list(
  transmute(df, x=Income, y=Limit, dataset=1),
  transmute(df, x=Income, y=Rating, dataset=2),
  transmute(df, x=Income, y=Age, dataset=3),
  transmute(df, x=Income, y=Balance, dataset=4),
  transmute(df, x=Income, y=bal_to_limit_ratio, dataset=5),
  transmute(df, x=Limit, y=Rating, dataset=6),
  transmute(df, x=Limit, y=Age, dataset=7),
  transmute(df, x=Limit, y=Balance, dataset=8),
  transmute(df, x=Limit, y=bal_to_limit_ratio, dataset=9),
  transmute(df, x=Rating, y=Age, dataset=10),
  transmute(df, x=Rating, y=Balance, dataset=11),
  transmute(df, x=Rating, y=bal_to_limit_ratio, dataset=12),
  transmute(df, x=Age, y=Balance, dataset=13),
  transmute(df, x=Age, y=bal_to_limit_ratio, dataset=14),
  transmute(df, x=Balance, y=bal_to_limit_ratio, dataset=15)
) %>%
bind_rows() %>%
ggplot(aes(x, y)) +
geom_point() + facet_wrap(~ dataset)
```


```{r}
library(corrplot)
corrplot(cor(df.new), method = "ellipse")
```
From here we can say that there is positive association between (x=Income, y=Limit),
   (  x=Income, y=Rating),
   (  x=Income, y=Age),
   (  x=Income, y=Balance),
   
   (  x=Limit, y=Rating),
   (  x=Limit, y=Age),
   (  x=Limit, y=Balance),
   (  x=Limit, y=bal_to_limit_ratio),
   (  x=Rating, y=Age),
   (  x=Rating, y=Balance),
   (  x=Rating, y=bal_to_limit_ratio),
   (  x=Age, y=Balance,  =13),
  
   (  x=Balance, y=bal_to_limit_ratio). Negative association between (  x=Income, y=bal_to_limit_ratio), (  x=Age, y=bal_to_limit_ratio),(education,cards) and almost no association between (cards, rating),(cards,age),(age,education) etc.
```{r}
library(dplyr)
library(ggplot2)
dfnew=df %>%
  dplyr:: select(Income,Age) %>%
  gather()
dfnew
df7=dfnew %>%
  ggplot( mapping=aes(x=value))+
  geom_histogram(color="blue",fill="springgreen3",bins=25) +scale_x_continuous(lim = c(20, 32))
df7+ facet_wrap(~key)+labs(x = "Magnitude")+ggtitle("New Faceted Histogram")
# df %>%
#   ggplot(aes(x=Cost))+
#   geom_histogram(color="blue",fill="springgreen3")
```

```{r}
library(dplyr)
library(ggplot2)
dfnew=df %>%
  dplyr:: select(bal_to_limit_ratio) %>%
  gather()
dfnew %>%
  ggplot( mapping=aes(x=value))+
  geom_histogram(color="blue",fill="springgreen3",bins=25) +scale_x_continuous(lim = c(0, 1))
```

```{r}
library(dplyr)
library(ggplot2)
dfnew=df %>%
  dplyr:: select(Limit) %>%
  gather()
dfnew %>%
  ggplot( mapping=aes(x=value))+
  geom_histogram(color="blue",fill="springgreen3",bins=25) +scale_x_continuous(lim = c(0, 10000))
```
```{r}
library(dplyr)
library(ggplot2)
dfnew=df %>%
  dplyr:: select(Education) %>%
  gather()
dfnew %>%
  ggplot( mapping=aes(x=value))+
  geom_histogram(color="blue",fill="springgreen3",bins=25) +scale_x_continuous(lim = c(0, 20))
```
These are histogram representation of Question 2.
```{r}
boxplot=ggplot(data=df,mapping=aes(x=Married,y=Balance)) + 
  geom_boxplot(fill="tomato2")
boxplot
library(GGally)
library(ggplot2)
ggpairs(df.new)
```
* * *
Some box plot interpretetion.
